%% Select the dissertation mode on
% See the documentation for more information about the available class options
% If you give option 'draft' or 'draft*', the draft mode is set on
%\documentclass[dissertation,final,nocontribution]{aaltoseries}

\documentclass[dissertation,final]{aaltoseries} %draft*->final
%\documentclass[dissertation]{aaltoseries}
\usepackage[utf8]{inputenc}

% Set the document languages
\usepackage[english]{babel}        % British hyphenation patterns

%\usepackage[ELEC, RGB]{aaltologo}

\usepackage[babel=true]{microtype} % http://ctan.org/tex-archive/macros/latex/contrib/microtype

\usepackage{graphicx}
 \graphicspath{{graphs/}}
 \DeclareGraphicsExtensions{.pdf}
 
\usepackage{amsmath}% http://ctan.org/pkg/amsmath
%%,amsfonts,amsthm,amsxtra,amsbsy,amssymb,mathrsfs}   % Mathematics
%\usepackage{bm} % bold math argument, http://www.ctan.org/pkg/bm

\usepackage{url}                   
                % http://ctan.org/tex-archive/macros/latex/contrib/url
\usepackage[caption=false,font=footnotesize]{subfig}                
                % http://ctan.org/tex-archive/macros/latex/contrib/subfig
\usepackage[labelfont=bf]{caption}  % , justification=centering
\usepackage{acronym}               
                % http://ctan.org/tex-archive/macros/latex/contrib/acronym
\usepackage{booktabs}              
                % http://ctan.org/tex-archive/macros/latex/contrib/booktabs
\usepackage{bytefield}             
                % http://ctan.org/tex-archive/macros/latex/contrib/bytefield
\usepackage{listings}              
                % http://ctan.org/tex-archive/macros/latex/contrib/listings
\usepackage{float}                 % Needed for hyperref - see note above
\usepackage{hyperref}              % Problematic - see note above
\usepackage{algorithm}             
                % http://ctan.org/tex-archive/macros/latex/contrib/algorithms
\usepackage{algorithmic}           % (also part of the algorithms bundle)

\usepackage{longtable}             % for tables that span across pages

\usepackage[footnote,draft,silent,nomargin]{fixme} %added FIXME!

\usepackage{multirow}   % for multirows
\usepackage{rotating}

\usepackage{setspace} 

%\floatstyle{plain}
%\newfloat{program}{thp}{figures}[chapter]
%\floatname{program}{Program}

\usepackage{chngcntr}
\counterwithout{footnote}{chapter}


\hyphenation{tele-presence}
\hyphenation{multi-media}

\hypersetup{
  pdfauthor={Varun Singh},
  pdftitle={Protocols and Algorithms for Adaptive Multimedia Systems},
  pdfsubject={RTP Congestion Control},
  pdfkeywords={RTP} {Congestion Control} {MPRTP} {Geo-location} 
              {Multipath RTP} {Interactive Multimedia}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% The author of the dissertation
\author{Varun Singh}
% The title of the thesis: Algorithms and Protocol extensions for Enabling Adaptive Multimedia Applications
\title{Protocols and Algorithms for Adaptive Multimedia Applications}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% The abstract of the dissertation in English
% Use this command!
\draftabstract{

The deployment of WebRTC and telepresence systems is going to enable wide-scale 
adoption of high quality real-time communication. Delivering high
quality video usually corresponds to an increase in required network capacity
and also requires an assurance of network stability. A real-time multimedia
application that uses the Real-time Transport Protocol (RTP) over UDP needs to
implement congestion control since UDP does not implement any such mechanism.
This thesis is about enabling congestion control for real-time communication,
and deploying it on the public Internet, which consists of a mixture of wired and
wireless links.

A congestion control algorithm relies on congestion cues, such as RTT and
loss. Hence, in this thesis, we first propose a framework for classifying
congestion cues. We classify the congestion cues as a combination of: where
they are measured or observed? And, how is the sending endpoint notified? For
each, there are two options, i.e., the cues are either observed and reported by
an on-path or by an off-path source, and, the cue is either reported in-band
or out-of-band, which results in four combinations. Hence, the framework
provides options to look at congestion cues beyond those reported by the
receiver.

We propose a sender-driven, a receiver-driven, and a hybrid congestion control
algorithm. The hybrid algorithm relies on both the sender and receiver 
co-operating to perform congestion control. Lastly, we compare the performance of
these different algorithms. We also explore the idea of using capacity
notifications from middleboxes (e.g., 3G/LTE base stations) along the path as
cues for a congestion control algorithm. Further, we observe the interaction
between error-resilience mechanisms and show that FEC can be used in a
congestion control algorithm for probing for additional capacity.

We propose Multipath RTP (MPRTP), an extension to RTP, which uses multiple
paths for either aggregating capacity or for increasing error-resilience. We
show that our proposed scheduling algorithm works in diverse scenarios (e.g.,
3G and WLAN, 3G and 3G, etc.) with paths with varying latencies.

Lastly, we propose a network coverage map service (NCMS), which aggregates
throughput measurements from mobile users consuming multimedia services. The
NCMS sends notifications to its subscribers about the upcoming network
conditions, which take these notifications into account when performing
congestion control.

In order to test and refine the ideas presented in this thesis, we have
implemented them in proof-of-concept prototypes, and conducted
experiments and simulations to validate our assumptions and gain new insights.
 

}
% Let's add another one in Finnish
%\draftabstract[finnish]{\lipsum[4-6]}
% And yet another one in Swedish
%\draftabstract[swedish]{\lipsum[7-9]}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Preface
% If you write this somewhere else than in Helsinki, use the optional location.
\begin{preface}[Espoo]

This dissertation is based on several years of research at Aalto University
(earlier Helsinki University of Technology) and standardisation at the
Internet Engineering Task Force (IETF). My sincere thanks to all the people who
helped me with the research. Foremost, I would like to thank Prof. J\"{o}rg Ott
for encouraging me to pursue the interaction of multimedia and congestion
control in more detail. He is an outstanding guide, and an inspiration. His 
timely interactions kept spirits high especially when faced with 
disappointments and for his patience to stay the course.

Igor Curcio and Nokia Research Center, presented the practical problems faced
by multimedia in mobile networks; his experience with mobile systems was
influential in shaping my early ideas related to congestion control.  
%
Colin Perkins, his book on RTP is a practitioners guide -- covers aspects of the
design which are not covered in the RFCs. Additionally, the numerous 
discussions I had with him during over the span of the last few years has been 
very beneficial. 
%
Lars Eggert encouraged me to contribute to the IETF and that participation 
has generated several new collaborations, including working with some of the
brightest minds in the Internet/networking world.

During the course of my research, I worked with several colleagues on a 
day-to-day basis. They were: Sharmistha Chatterjee (2010), Saba Ahsan (2011), 
Marcin Nagy (2012), and Albert Abello (2013). This experience was tremendously 
beneficial because I learnt several skills beyond just performing research.

An equal measure of gratitude is due to all the co-authors of  the papers that
make up the thesis. They are: J\"{o}rg Ott, Igor Curcio, Saba Ahsan, Marcin 
Nagy, Albert Abello, Colin Perkins, Lars Eggert, Jegadish Devadoss,  Martin  
Ellis, Chenghao Liu, and Ye-Kui Wang.

The pre-examiners for this thesis were Professors Carsten Griwodz and J\"{o}rg
Widmer. I want to thank them both for their efforts in reviewing the thesis. 
Their comments and suggestions helped improve the quality of the thesis. 

Special thanks my friends: Irena, L\'{a}szl\'{o}, M\'{a}rio, Cathy, Jakub, 
Markus, Jos\'{e}-Luis, Pavan and Tatu for the help and support during
the years of research. An extended gratitude to the folks at Aalto 
Entrepreneurship Society and Startup Sauna -- you rock! To my family for
the sacrifice they have patiently made. Finally, I thank Irena for helping
me understand that the benefits out-weigh the risk when following your heart.

\end{preface}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Table of contents of the dissertation
\tableofcontents

% For article dissertations, remove if you write a monograph dissertation.
\publicationdataorder{author}{title}{publication}{information}{month}{year and page}
\publicationpunctuation[]{,}{}{,}{,}{,}{}
\publicationformatting[]{}{}{\em}{}{}{}
\publicationadditionalpreformatting{}{``}{}{}{}{}
\publicationadditionalpostformatting{}{,''}{}{}{}{}

\listofpublications

%\listoffixmes

%% Add lists of figures and tables as you usually.
 \listoffigures
 \listoftables
% \listofalgorithms
% \listofequations

%% Add list of abbreviations, list of symbols, etc., using your preferred
%% package/method.

%\include{listofabbreviations}
%\include{listofsymbols}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter*{List of Abbreviations}
\input{./chapters/_0lists}

\chapter*{List of Symbols}
\input{./chapters/_0symbols}

%% The main matter, one can obviously use \input or \include
\chapter{Introduction}
\label{chap1}
\input{./chapters/_1intro}

% \chapter{Research Goal}
% \label{chap:rg}
% \input{./chapters/_2rg}

\chapter{RTP: Real-time Transport Protocol}
\label{chap:rtp}
\input{./chapters/_3rtp}

\chapter{Congestion Control Framework for Real-time Communication}
\label{chap:cc.fw}
\input{./chapters/_4fw}

\chapter{Congestion Control for Interactive Multimedia}
\label{chap:cc}
\input{./chapters/_5rtc}
%% An example for changing the running header (the optional parameter)

 \chapter{Interaction of Error-Resilience and Congestion Control}
 \label{chap:er-cc}
\input{./chapters/_6ercc}

\chapter{Mobility, Offloading, Multihoming, and Overlays}
\label{chap:mprtp}
\input{./chapters/_7mprtp}

 \chapter{Network-assisted Congestion Control}
 \label{chap:cc.nw}
 \input{./chapters/_8nwc}

\chapter{Conclusions}
\label{chap:conc}
 \input{./chapters/_9conc}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% The following commands are for article dissertations, remove them if you
%% write a monograph dissertation.

% Errata list, if you have errors in the publications.
%\errata

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\bibliographystyle{IEEEtran} %plain, amsalpha

\bibliographystyle{plain}

% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../allpapers}

\bibliography{bib/diss,bib/rfc}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% The first publication (journal article)
% Set the publication information.
% This command musts to be the first!

\addpublication[conference]{V. Singh, S. McQuistin, M. Ellis, C. Perkins}
{Circuit Breakers for Multimedia Congestion Control}
{Proceedings of IEEE International Packet Video Workshop}
{San Jose, USA}{Dec}{2013, Pages 1--8}{IEEE}{c:cb}

% Add the dissertation author's contribution to that publication.

\addcontribution{The author of this dissertation was the main author of 
this paper and co-edited the paper with Colin Perkins. He equally contributed 
to the ideas and concepts discussed in the paper.  He also implemented the 
algorithm, conducted and analysed experiments for the interactive video 
scenario.}

% Add the errata of the publication, remove if there are none (the order can
% be interchanged with \addauthorscontribution).
%\adderrata{This is wrong}
% Add the publication pdf file, the filename is the parameter (must be the
% last).

\addpublicationpdf{add/9_cb.pdf}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% The second publication (conference article, note the optional parameter)
% Set the publication information.

\addpublication[conference]{V.Singh, J. Ott, I.D.D. Curcio}
{Rate adaptation for 3G Conversational Video}
{Proceedings of IEEE Infocom Workshops}
{Rio de Janeiro, Brazil}{Apr}{2009, Pages 1--7}{IEEE}{c:3grc}

% Add the dissertation author's contribution to that publication.

\addcontribution{The author of this dissertation was the main author of the
paper. He equally contributed to the ideas and concepts discussed in the paper,
and also implemented the algorithm and the overall system, conducted the
experiments and analysed the results.}

\addpublicationpdf{add/2_3g_ra.pdf}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% The third publication (another journal paper, accepted for publication,
%% note the optional parameter)
% Set the publication information, detailed information can be empty

\addpublication[conference]{V. Singh, J. Ott, I.D.D. Curcio}
{Rate Adaption for Conversational Video in Heterogeneous Environments}
{Proceedings of IEEE International Symposium on a World of Wireless, Mobile and Multimedia (WoWMoM)}
{San Francisco, USA}{Jun}{2012, Pages 1--7}{IEEE}{c:hetrc}

% Add the dissertation author's contribution to that publication.

\addcontribution{The author of this dissertation was the main author of the
paper. He equally contributed to the ideas and concepts discussed in the 
paper, and implemented the algorithm and the overall system, conducted the 
experiments and analysed the results.}

\addpublicationpdf{add/3_het_ra.pdf}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\addpublication[conference]{V. Singh, A. A. Lozano, J. Ott}
{Performance Analysis of Receive-Side Real-Time Congestion Control for WebRTC}
{Proceedings of IEEE International Packet Video Workshop}
{San Jose, USA}{Dec}{2013, Pages 1--8}{IEEE}{c:eval}

% Add the dissertation author's contribution to that publication.

\addcontribution{The author of this dissertation was the main author of this
paper. His contribution consisted of designing the experiments and analysing the results discussed in the paper.}

\addpublicationpdf{add/10_rrtcc.pdf}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\addpublication[conference]{J. Devadoss, V. Singh, C. Liu, J. Ott, Y-K. Wang,
I.D.D. Curcio}{Evaluation of Error Resilience Mechanisms for 3G Conversational
Video}{Proceedings of IEEE International Symposium on Multimedia (ISM)}
{Berkeley, USA}{Dec}{2008, Pages 378--383}{IEEE}{c:err}

% Add the dissertation author's contribution to that publication (the order
% can be interchanged with \adderrata).

\addcontribution{The author of this dissertation was one of the co-authors of
the paper. His contribution consisted of providing the 2 out of the 4 ideas
(NACK and Slice-size adaptation) discussed in the paper, implementing the
corresponding algorithms, conducting the  experiments and  co-editing the paper
with the main author of the paper (Jegadish Devadoss).}

\addpublicationpdf{add/1_3g_er.pdf}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\addpublication[submitted]{M. Nagy, V. Singh, J. Ott, L. Eggert}
{Rate-control using FEC for Interactive Multimedia Communication}
{Proceedings of ACM Multimedia Systems (MMSys)}
{Singapore, Singapore}{March}{2014, Pages 191--202}{ACM}{c:fecrc}

% Add the dissertation author's contribution to that publication.

\addcontribution{The author of this dissertation was one of the co-authors of
the paper. His contribution consisted of discussing the main idea and concept
for the paper with the lead author, designing the experiments, implementing the
comparative algorithms and providing the corresponding results, and co-editing
the paper with the main author of the paper (Marcin Nagy).}

\addpublicationpdf{add/6_fec_ra.pdf}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\addpublication[conference]{V. Singh, S. Ahsan, J. Ott}
{MPRTP: Multipath Considerations for Real-time Media}
{Proceedings of ACM Multimedia Systems (MMSys)}
{Oslo, Norway}{Feb}{2013, Pages 190--201}{ACM}{c:mprtp}

% Add the dissertation author's contribution to that publication.

\addcontribution{The author of this dissertation was the main author of the
paper. His contribution consisted of providing the main
idea for the paper, providing the configurations for the experiments,
performing the statistical analysis, and acting as the lead editor of the
paper.}

\addpublicationpdf{add/5_mprtp.pdf}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% The fourth publication (yet another journal paper, submitted for
%% publication, note the optional parameter) Note that you are allowed to use
%% this option only when submitting the dissertation for pre-examination!
% Set the publication information, detailed information is not printed

\addpublication[conference]{V. Singh, J. Ott, I.D.D. Curcio}
{Predictive Buffering for Streaming Video in 3G Networks}
{Proceedings of IEEE International Symposium on a World of Wireless, Mobile and Multimedia (WoWMoM)}
{San Francisco, USA} {Jun}{2012, Pages 1--10}{IEEE}{c:glass}

% Add the dissertation author's contribution to that publication.

\addcontribution{The author of this dissertation was the main author of the
paper. He equally contributed to the ideas and concepts discussed in the paper,
he implemented the algorithm and the overall system, conducted the experiments
and analysed the results.}

\addpublicationpdf{add/4_glass.pdf}

\end{document}
